---
title: "Glycimic Index"
author: "Richel Bilderbeek"
date: "10/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r}
library(ggplot2)
library(dplyr)
```

## Loading data

Control, with white bread, which has a known glycemic index of 75%:

```{r}
t_control <- read.csv("control.csv")
names(t_control) <- c("t", names(t_control)[-1])
t_control$Gemiddeld <- NULL
knitr::kable(t_control)
```

Treatment, with Jimmy Joy's Plenny Shake, of which the glycemic
index needs to be determined:

```{r}
t_jj <- read.csv("jj.csv")
names(t_jj) <- c("t", names(t_jj)[-1])
t_jj$Gemiddeld <- NULL
knitr::kable(t_jj)
```

## Convert data to long form

```{r}
t_control_long <- tidyr::pivot_longer(
  t_control,
  cols = starts_with("X"),
  names_to = "person_id"
)
t_control_long$treatment <- "control"

t_jj_long <- tidyr::pivot_longer(
  t_jj,
  cols = starts_with("X"),
  names_to = "person_id"
)
t_jj_long$treatment <- "jj"
t <- rbind(t_control_long, t_jj_long)
knitr::kable(head(t))
```

## AUC per person in time

```{r}
t$person_id <- as.factor(t$person_id)
ggplot(
  t,
  aes(x = t, y = value, color = treatment)
) + geom_point() + geom_smooth() +
  scale_y_continuous("AUC of blood glucose concentration") +
  scale_x_continuous("time (minutes)") +
  ggtitle("AUCs in time") +
  facet_grid(person_id ~ .) +
  ggsave("aucs_in_time_pp.png", width = 7, height = 49)
```

## Average AUC in time

```{r}
ggplot(
  t,
  aes(x = t, y = value, color = treatment)
) + geom_point() + geom_smooth() +
  scale_y_continuous("AUC of blood glucose concentration") +
  scale_x_continuous("time (minutes)") +
  ggtitle("AUCs in time") +
  ggsave("aucs_in_time.png", width = 7, height = 7)
```

## Calculate the glycemic index

### Calculate the average concentrations (again)

```{r}
t_summary <- t %>% dplyr::group_by(t, treatment) %>% dplyr::summarise(mean = mean(value), .groups = "drop")
knitr::kable(t_summary)
```

### Approximate the area in the two-hour curve

Simply sum the values

```{r}
t_gi <- t_summary %>% dplyr::group_by(treatment) %>% dplyr::summarise(sum = sum(mean), .groups = "drop")
knitr::kable(head(t_gi))
```
### Correct values

We know the control (white bread) has a glycemic index of 75%.

```{r}
control_row <- which(t_gi$treatment == "control")
control_sum <- t_gi$sum[control_row]
conversion <- 75 / control_sum
testthat::expect_equal(control_sum * conversion, 75)
jj_row <- which(t_gi$treatment == "jj")
jj_sum <- t_gi$sum[jj_row]
gi <- jj_sum * conversion
cat(paste("The glycemic index of Jimmy Joy's Plenny shake is", gi))
```


The glycemic index of Jimmy Joy's Plenny shake is approximately 40%,
which is classified as a low glycemic index.
